---
title: "4180_CPSC_Midterm_Project1"
author: "Yatri Patel"
date: "9/30/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# CPSC 4180 - Midterm Project 1 

In this exam, you are asked to analyze the Global COVID19 time-series data at https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series (Links to an external site.) You are expected to generate a figure that overlays the COVID19 daily cases in a country with the average Google Global Mobility data available at:  https://www.google.com/covid19/mobility/ (Links to an external site.). You are recommended to average the regional mobility values and daily cases in that country in order to overlay with this COVID19 cases at the entire country. For undergraduates,  you are expected to investigate the COVID19 daily cases and mobility in one country,  either Italy, UK, or India. 

# Importing libraries needed for the project

```{r}
rm(list = ls()) #clean up the work space
library(tidyverse)
library(lubridate)
library(reshape2)
library(ggplot2)
library(dplyr)
```

# Load the covid-19 data 

```{r}
#link to the covid global data
#covid_global <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
#download the covid data 
#download.file(covid_global, 'time_series_covid19_confirmed_global.csv')
#read the downloaded file
tb = read_csv('time_series_covid19_confirmed_global.csv')
```

# Italy Daily COVID Cases data  

```{r}
#assign my country variable to country of choice
my_country <- "Italy"
```

```{r}
tb1 <- tb %>% dplyr::select(12:length(tb[1,])) %>% as.matrix()
tb1 <- as.data.frame(t(tb1))

head(tb1)

names(tb1) = tb$`Country/Region`

current_dates = names(tb)[12:length(tb[1,])]
tb1$dates = mdy(current_dates)
head(tb1)
```

# Get daily confirmed covid cases data for the country of choice

```{r}
#number of the days
days_window = 609
endday = today(tzone = 'CET') #define time zone

tb_italy <- tb1 %>%
  dplyr::select(all_of(my_country), 'dates') %>%
  dplyr::filter(between(dates, endday-ddays(days_window), endday))

head(tb_italy)
```

```{r}
#calculate the daily covid cases of italy
local_daily_italy = c(0, diff(tb_italy[,1]))

tb_daily_italy = data.frame(local_daily_italy)

head(tb_daily_italy)
```

```{r}
tb_daily_italy$YMD = mdy(row.names(tb_italy))
head(tb_daily_italy)
```

```{r}
names(tb_daily_italy)[1] = my_country 
```

# Plotting the daily covid-19 cases

```{r}
my_plot <-  ggplot(tb_daily_italy, aes(x=YMD, y = tb_daily_italy[,1]))
my_plot + geom_point() + stat_smooth(span= 0.3) + ggtitle("Daily COVID-19 Cases in Italy") +ylab("Daily cases") + xlab("Time")
```

# Mobility Data

```{r}
#read mobility data csv
tbGM = read_csv('Global_Mobility_Report.csv')
head(tbGM)
```

```{r}
#select the mobility data of the country
tbGM_italy <- tbGM %>% filter(country_region == my_country)
head(tbGM_italy)
```

```{r}
#max date in the mobility data
max(tbGM_italy$date)
```

```{r}
head(tb_daily_italy)
```

```{r}
#make a copy of the daily italy covid cases data 
tb_daily_italy_copy = tb_daily_italy
names(tb_daily_italy_copy) = c('DailyCases', 'date')
head(tb_daily_italy_copy)
```

```{r}
#merge the mobility and the covid data
tb_GMCovid_Italy = merge(x=tb_daily_italy_copy, y=tbGM_italy, by = 'date', all.y = TRUE)

#remove the date column 
tb_GMCovid_Italy2 <- tb_GMCovid_Italy %>% dplyr::select(2, 11:16)
head(tb_GMCovid_Italy2)
```

```{r}
#scale the data set 
tb_GMCovid_Italy_scaled <- data.frame(scale(tb_GMCovid_Italy2))

head(tb_GMCovid_Italy_scaled)
```

```{r}
#add the date to the scaled data
tb_GMCovid_Italy_scaled$date = tb_GMCovid_Italy$date
head(tb_GMCovid_Italy_scaled)
```

# Plot the overlayed map 

```{r}
#select the columns
selected_columns = c('DailyCases', 'workplaces_percent_change_from_baseline', 'residential_percent_change_from_baseline')

#melt the two dataframes
df_melt <- melt (tb_GMCovid_Italy_scaled, measure.vars = selected_columns, value.name = "Value", variable.name= "variable")

#plot the melted dataframe
plot <- ggplot(df_melt, aes(x=date, y=df_melt$Value, color=variable)) + stat_smooth(span= 0.3)

#add title and x and y label 
plot +ggtitle("Overlay plot of mobility and covid cases data") + ylab("Value") + xlab("Date")
```


